---
title: "Lab08: Breast Cancer Mini Project"
author: "Minzhou Ou (PID: A17293812)"
format: pdf
---

## Background

In today's class we will be employing all the R techniques for data analysis that we have learned thus far - including the machine learning methods of clustering and PCA - to analyze real breast cancer biopsy data. 

## Data Import

The data is in CSV format:

```{r}
WI_cancer <- read.csv("WisconsinCancer.csv", row.names=1)
```

wee peak at the data
```{r}
head(WI_cancer,3)
WI_cancer.data <- WI_cancer[, -1]
```

> Q1. How many observations are in this dataset?

```{r}
nrow(WI_cancer)
```

> Q2. How many of the observations have a malignant diagnosis?

```{r}
sum(WI_cancer$diagnosis == "M")
tbl <- table(WI_cancer$diagnosis)
pct <- tbl / sum(tbl) * 100
tbl
pct
```

> Q3. How many variables/features in the data are suffixed with _mean?

```{r}
length(grep("_mean", colnames(WI_cancer)))
```

We need to remove the `diagnosis` column before we do any further analysis of this dataset - we don't want to pass this to PCA etc. We will save it as a separate wee vector that we ca use later to compare our findings to those of experts

```{r}
unsupervised <- WI_cancer[ ,-1]
diagnosis <- WI_cancer$diagnosis
```

## Principal Component Analysis (PCA)

The main function in base R is called `prcomp()` we will use the optional argument `scale=TRUE` here as the data columns/features/dimensions are on very different scales in the original data set

```{r}
WI_cancer.pr <- prcomp(unsupervised, scale=T)

```

> Q4. From your results, what proportion of the original variance is captured by the first principal component (PC1)?

```{r}
summary(WI_cancer.pr)
```
The proportion of the PC1 varaince is 44.27%

> Q5. How many principal components (PCs) are required to describe at least 70% of the original variance in the data?

We need at least PC3

> Q6. How many principal components (PCs) are required to describe at least 90% of the original variance in the data?

We need PC7 to get 90%

> Q7. What stands out to you about this plot? Is it easy or difficult to understand? Why?

```{r}
biplot(WI_cancer.pr)
```

The data looks messy and hard to follow, because the values are not normalized, and there's too much data doesn't show trend or anything

```{r}
library(ggplot2)

ggplot(WI_cancer.pr$x) +
  aes(PC1, PC2, col=diagnosis) +
  geom_point()
```

## Variance

```{r}
pr.var <- WI_cancer.pr$sdev^2
head(pr.var)
pve <- pr.var / sum(pr.var)
plot(c(1,pve), xlab = "Principal Component", 
     ylab = "Proportion of Variance Explained", 
     ylim = c(0, 1), type = "o")
barplot(pve, ylab = "Percent of Variance Explained",
     names.arg=paste0("PC",1:length(pve)), las=2, axes = FALSE)
axis(2, at=pve, labels=round(pve,2)*100 )
```

> Q9. For the first principal component, what is the component of the loading vector (i.e. wisc.pr$rotation[,1]) for the feature concave.points_mean? This tells us how much this original feature contributes to the first PC. Are there any features with larger contributions than this one?

```{r}
pc1 <- WI_cancer.pr$rotation[,"PC1"]
WI_cancer.pr$rotation["concave.points_mean", 1]
min(pc1)
```


## 4. Hierarchical clustering

```{r}
data.scaled <- scale(unsupervised)
data.dist <- dist(data.scaled)
wisc.hclust <- hclust(data.dist, method = "ward.D2")
plot(wisc.hclust)
abline(wisc.hclust, col="red", lty=2)
```

## Combining methods

The idea here is taht I can take my new variables (the PCs) that are better descriptors of the data-set than the original features (i.e. the 30 columns in `wisc.data`) and use these as a basis of clustering. 

```{r}
pc.dist <- dist(WI_cancer.pr$x[ ,1:3])
wisc.pr.hclust <- hclust(pc.dist, method = "ward.D2")
plot(wisc.pr.hclust)
grps <- cutree(wisc.pr.hclust, k=2)
table(grps)
```
I can now run `table()` with both myu clustering `grps` and the expert `diagnosis`

```{r}
table(grps, diagnosis)
```

Our cluster "1" has 179 "M" diagnosis
or cluster "2" has 333 "B" diagnosis

179 True Positive
24 False Positive
333 True Negative
33 False negative

Sensitivity: TP/(TP+FN)
```{r}
sensitivity <- 179/sum(179,33)
sensitivity
```
Specificity: TN/(TN+FP)
```{r}
specificity <- 333/sum(333,24)
specificity
```

## Prediction

We can use our PCA model for prediction of new un-seen data

```{r}
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(WI_cancer.pr, newdata=new)
npc
plot(WI_cancer.pr$x[,1:2], col=grps)
points(npc[,1], npc[,2], col="blue", pch=16, cex=3)
text(npc[,1], npc[,2], c(1,2), col="white")
```


