---
title: "Lab06: R Functions"
author: "Minzhou Ou (PID: A17293812)"
format: pdf
---

## Backgroound

Functions are at the heart of using R. Everything we do involves calling and using functions (from data input, analysis to results)

All functions in R have at least 3 things:

1. A **name** the thing we use to call the function.
2. One or more input **arguments** that are comma separated
3. The **body**, lines of code between curly `{ }` that does the work of the function

## A first function

Let's write a silly wee function to add some number:

```{r}
add <- function(x) {
  x + 1
}
```

Let's try it out

```{r}
add(100)
```

Will this work

```{r}
add(c(100, 200, 300))
```

Modify to be more useful and add more than just 1

```{r}
add <- function(x, y) {
  x + y
}
```

```{r}
add(100, 10)
```

Will this work?

```{r}
#add(199)
```

```{r}
plot(1: 10, col = "blue", typ = "b")
```

```{r}
log(10, base = 10)
```

> **N.B.** input argument can be **required** or **optional**. The later have a fall-back default that is specifed in the function code with an equals sign. 

```{r}
#add(x = 100, y = 200, z = 300)
```

## A second function

All functions in R look like this

```
name <- function(arg) {
  body
}
```

The `sample()` function in R ...

```{r}
sample(1:10, size = 4)
```

> Q. Returnn 12 numbers picked ranndomly from the input 1: 10

```{r}
sample(1:10, size = 12, replace = TRUE)
```

> Q. Write the code to generate a random 12 nucleotide long DNA sequence

```{r}
sample(c('A','C','G','T'), size = 12, replace = TRUE)
```

> Q. Write a first versionn function called `generate_dna()` that generates a user specified lenngth `n` random DNA sequence. It is compact

```{r}
generate_dna <- function(n) {
  nucleotide <- c('A','C','G','T')
  DNA <- sample(nucleotide, size = n, replace = TRUE)
  paste(DNA, collapse="")
}
generate_dna(pi)
```

```{r}
# Example pattern (not using your bases)
x <- c("H", "E", "L", "L", "O")
paste(x, collapse = "hello")
# returns "HELLO"
```

> Q. Give the user an option to return FASTA format output sequence or standard multi-elemment vector format?

```{r}
generate_dna <- function(n=6, fasta=TRUE) {
  nucleotide <- c('A','C','G','T')
  DNA <- sample(nucleotide, size = n, replace = TRUE)
  if(fasta) {
    DNA <- paste(DNA, collapse="")
    cat("Hello...")
  }
  else{
    cat("...from the other side...")
  }
  return(DNA)
}
```
```{r}
generate_dna(10)
generate_dna(10, FALSE)
```

## A new cool function

> Q. Write a function called `generate_protein()` that generates a user specifed length protein sequence in FASTA like format

```{r}
generate_protein <- function(length){
  amino_acides <- c("A","R","N","D","C","E","Q","G","H","I","L","K","M","F","P","S","T","W","Y","V")
  proteins <- sample(amino_acides, size = length, replace = TRUE)
  proteins <- paste(proteins, collapse="")
  return(proteins)
}

generate_protein(10)
```


> Q. Use your new `generate_protein()` function to generate sequences between length 6 and 12 amino-acides in length and check of any of these are unique in nature (i.e. found in the NR database at NCBI)

```{r}
library(rentrez)
```

```{r}
generate_protein <- function(){
  amino_acides <- c("A","R","N","D","C","E","Q","G","H","I","L","K","M","F","P","S","T","W","Y","V")
  length = sample(6:12, replace = TRUE)
  proteins <- sample(amino_acides, size = length, replace = TRUE)
  proteins <- paste(proteins, collapse="")
  return(proteins)
}
generate_protein()
```
```{r}
generate_protein <- function(length){
  amino_acides <- c("A","R","N","D","C","E","Q","G","H","I","L","K","M","F","P","S","T","W","Y","V")
  # length = sample(6:12, replace = TRUE)
  proteins <- sample(amino_acides, size = length, replace = TRUE)
  proteins <- paste(proteins, collapse="")
  return(proteins)
}
for(i in 6: 12){
  cat(">", i, "\n", sep = "")
  cat(generate_protein(i), "\n")
}
```
They can be blast



```{r}
library(bio3d)
protein_drug_iteraction <- function(pdi_seq, chain, elety){
  pdi_sequence <- read.pdb("pdi_seq")
  seq_trim <- trim.pdb(pdi_sequence, chain=chain, elety=elety)
  seq_b <- seq_trim$atom$b
  plotb3(seq_b, sse=seq_trim, typ="l", ylab="Bfactor")
  return(seq_trim)
} 
```

